Location and Scale-Invariant Power Transformations for Transforming Data to Normality

### 2024-08-21 manuscript/manuscript.Rmd
Additional changes.

There are a few options based on which weights may be set. Here, we investigate three:

@@

Based on either of the above, weights are set using a weighting function. We will investigate the following functions, where $\dot{x}_i$ can be either $p^{*}_i=2 \left( p_i - 0.5\right)$ (so that the quantiles are zero-centred), $z_i$ or $r_i$:

@@

```{=latex}
\begin{equation}
f_{\alpha}(x; \mu, \sigma, \beta) = \frac{2 \alpha \left(1 - \alpha\right)}{\sigma}
\begin{cases}
f_{\beta}\left( \left(1 - \alpha \right) \frac{\left| x - \mu \right|}{\sigma} \right) & \text{, } x \leq \mu \\
f_{\beta}\left( \alpha \frac{\left| x - \mu \right|}{\sigma} \right) & \text{, } x > \mu
\end{cases}
\end{equation}
```
Here $\alpha \in (0,1)$ is a skewness parameter. $\alpha > 0.5$ creates a distribution with a negative skew, i.e. a left-skewed distribution. A right-skewed distribution is created for $\alpha < 0.5$. $\mu \in \mathcal{R}$ and $\sigma \in (0, \infty)$ are location and scale parameters, respectively. $f_{\alpha}$ thus describes the probability density function of an asymmetric generalised normal distribution, which we will refer to here and parametrise as $\mathcal{AGN}\left(\mu, \sigma, \alpha, \beta \right)$.

@@

Power transformations aim to transform features to a normal distribution. However, this may not always be successful or possible. Deviations from normality can be detected by normality tests, such as the Shapiro-Wilk test [@Shapiro1965-zd]. In practice, normality tests may be too stringent to assess goodness of fit in presence of large sample sizes, outliers or a combination of the two. Here we develop an empirical test for central normality. The null hypothesis $H_0$ is that the distribution is centrally normal. The alternative hypothesis $H_1$ is that the distribution is not centrally normal.

@@

In this section we simulate data to assess invariance to location and scale of the proposed power transformations, weighting for robust transformations, and to develop the empirical central normality test.
The $\lambda$ parameter for conventional power transformations (Eqn. \ref{eqn:box-cox-original} and \ref{eqn:yeo-johnson-original}) and $\lambda$, $x_0$ and $s$ parameters for location and scale invariant power transformations (Eqn. \ref{eqn:box-cox-invariant} and \ref{eqn:yeo-johnson-invariant}) are estimated using the BOBYQA algorithm for derivative-free bound constraint optimisation [@Powell2009-zb] through maximum likelihood estimation. Appendix B describes other criteria, and Appendix C the results of simulations using these criteria.

@@

To assess whether the proposed power transformations lead to $\lambda$ values that are invariant to location and scale of the distribution, we simulate three different distributions. The first distribution is a normal distribution, as previously shown in Figure \ref{fig:decreased-normality}.
We first randomly draw $10000$ values from a normal distribution: $\mathbf{X}_{\text{normal}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{N}\left(0, 1\right)$, or equivalently $\mathbf{X}_{\text{normal}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{AGN}\left(0, 1/\sqrt{2}, 0.5, 2\right)$.
The second distribution is a right-skewed normal distribution $\mathbf{X}_{\text{right}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{AGN}\left(0, 1/\sqrt{2}, 0.2, 2\right)$.
The third distribution is a left-skewed normal distribution $\mathbf{X}_{\text{left}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{AGN}\left(0, 1/\sqrt{2}, 0.8, 2\right)$.
We then compute transformation parameter $\hat{\lambda}$ using the original definitions (equations \ref{eqn:box-cox-original} and \ref{eqn:yeo-johnson-original}) and the location and scale invariant definitions (equations \ref{eqn:box-cox-invariant} and \ref{eqn:yeo-johnson-invariant}) for each distribution.
To assess location invariance, a positive value $d_{\text{shift}}$ is added to each distribution with $d_{\text{shift}} \in [1, 10^6]$. Similarly, to assess scale invariance, each distribution is multiplied by a positive value $d_{\text{scale}}$, where $d_{\text{scale}} \in [1, 10^6]$.

>>>>

Weights may be set using several methods. Here, we investigate three:

@@

Weights are set using a weighting function for all three methods. We will investigate the following functions, where $\dot{x}_i$ can be either $p^{*}_i=2 \left( p_i - 0.5\right)$ (so that the quantiles are zero-centred), $z_i$ or $r_i$:

@@

```{=latex}
\begin{equation}
f_{\alpha}(x; \mu, \sigma, \beta) = \frac{2 \alpha \left(1 - \alpha\right)}{\sigma}
\begin{cases}
f_{\beta}\left( \left(1 - \alpha \right) \frac{\left| x - \mu \right|}{\sigma} \right) & \text{, } x \leq \mu \\
f_{\beta}\left( \alpha \frac{\left| x - \mu \right|}{\sigma} \right) & \text{, } x > \mu
\end{cases}
\end{equation}
```
Here, $\alpha \in (0,1)$ is a skewness parameter. $\alpha > 0.5$ creates a distribution with a negative skew, i.e. a left-skewed distribution. A right-skewed distribution is created for $\alpha < 0.5$. $\mu \in \mathcal{R}$ and $\sigma \in (0, \infty)$ are location and scale parameters, respectively. $f_{\alpha}$ thus describes the probability density function of an asymmetric generalised normal distribution, which we will refer to here and parametrise as $\mathcal{AGN}\left(\mu, \sigma, \alpha, \beta \right)$.

@@

Power transformations aim to transform features to a normal distribution. However, this may not always be successful or possible. Deviations from normality can be detected by normality tests, such as the Shapiro-Wilk test [@Shapiro1965-zd]. In practice, normality tests may be too stringent to assess goodness of fit in presence of large sample sizes, outliers, or both. Here we develop an empirical test for central normality. The null hypothesis $H_0$ is that the distribution is centrally normal. The alternative hypothesis $H_1$ is that the distribution is not centrally normal.

@@

In this section we simulate data to assess invariance to location and scale of the proposed power transformations, weighting for robust transformations, and to develop the empirical central normality test.
The $\lambda$ parameter for conventional power transformations (Eqn. \ref{eqn:box-cox-original} and \ref{eqn:yeo-johnson-original}), as well as $\lambda$, $x_0$ and $s$ parameters for location and scale invariant power transformations (Eqn. \ref{eqn:box-cox-invariant} and \ref{eqn:yeo-johnson-invariant}) are estimated using the BOBYQA algorithm for derivative-free bound constraint optimisation [@Powell2009-zb] through maximum likelihood estimation.

@@

To assess whether the proposed power transformations lead to values of $\lambda$ that are invariant to location and scale of the distribution, we simulate three different distributions. The first distribution is a normal distribution, as previously shown in Figure \ref{fig:decreased-normality}.
We first randomly draw $10000$ values from a normal distribution: $\mathbf{X}_{\text{normal}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{N}\left(0, 1\right)$, or equivalently $\mathbf{X}_{\text{normal}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{AGN}\left(0, 1/\sqrt{2}, 0.5, 2\right)$.
The second distribution is a right-skewed normal distribution $\mathbf{X}_{\text{right}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{AGN}\left(0, 1/\sqrt{2}, 0.2, 2\right)$.
The third distribution is a left-skewed normal distribution $\mathbf{X}_{\text{left}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{AGN}\left(0, 1/\sqrt{2}, 0.8, 2\right)$.
We then compute transformation parameter $\hat{\lambda}$ using the original definitions (equations \ref{eqn:box-cox-original} and \ref{eqn:yeo-johnson-original}) and the location and scale invariant definitions (equations \ref{eqn:box-cox-invariant} and \ref{eqn:yeo-johnson-invariant}) for each distribution.
To assess location invariance, a positive value $d_{\text{shift}}$ is added to each distribution with $d_{\text{shift}} \in [1, 10^6]$. Similarly, to assess scale invariance, each distribution is multiplied by a positive value $d_{\text{scale}}$, where $d_{\text{scale}} \in [1, 10^6]$.
